---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
keep_md: true
---

This document is for the Peer Assessment-1 of the reproducsible research. 

## Loading and preprocessing the data
```{r, echo=TRUE}
currentdir = getwd()
library(data.table)

#filepath = paste(currentdir, "/activity.csv", sep="")
filepath = "C:/Srini/Coursera/5-ReproducibleResearch/RepData_PeerAssessment1/data/activity.csv"
data <- read.csv("activity.csv")
aggregate(steps ~ date, data, sum)

# Remove the NA values
cleandata = na.omit(data)
tabledata <- data.table(data)
tabledata[, minutesseq := sequence(.N), by = c("date")]

```


## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day
```{r, echo=TRUE}
stepsperday = aggregate(steps ~ date, cleandata, sum)
steps = stepsperday$steps
```


2. Histogram of steps taken each day.

```{r, echo=TRUE}
hist(steps, col = c("blue", "red", "yellow", "green", "violet", "orange"))
```

2. Calculate and report the mean and median of the total number of steps taken per day

The mean number of steps taken per day is:
```{r, echo=TRUE}
mean(stepsperday$steps)
```

The median is:
```{r, echo=TRUE}
#mean steps per day
median(stepsperday$steps)
```


## What is the average daily activity pattern?
1. Make a time series plot (i.e. type = "l") of the 5-minute
To make the time series plot of 5 minute interval along the average number of steps take in the 5 minute interval, the
data is necessary for the avargae steps in the each of the 5 minute interval across all days.
```{r, echo=TRUE}
library(plyr)
stepsineachinterval = ddply(cleandata, .(interval), summarize,  avesteps=mean(steps))
plot(stepsineachinterval$interval, stepsineachinterval$avesteps, type = "l", col = "blue")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
Finding the 5 minute interval in which, at an averge, the maximum number of steps were taken.
```{r, echo=TRUE}
stepsineachinterval[which.max(stepsineachinterval$avesteps),]

```
The maximum number of steps, on average, are taken in the minute interval between 8:30 and 8:35 am.

Making a time series plot of 5-minute interval

## Imputing missing values

```{r, echo=TRUE}
# Updating the NA values with average steps in that interval.
# For example, if the observation at 810 is missing, it will be set to the avarage number of steps take in that interval.

# Copy the data to the new dataframe - dataupdated
dataupdated = data
totalrows = nrow(dataupdated)

count = 0
for (i in 1:nrow(dataupdated)) {
  if (is.na(dataupdated[i, 1])) {
    count = count + 1
    matchinginterval = dataupdated[i,3]
    newval = subset(stepsineachinterval, interval == matchinginterval)[1,2]
    dataupdated[i,1] = newval
  }
}


```


## Histogram with the dataset where NA value is replaced with the average value
```{r, echo=TRUE}
stepsperday = aggregate(steps ~ date, dataupdated, sum)
steps = stepsperday$steps
hist(steps, col = c("blue", "red", "yellow", "green", "violet", "orange"))

```
